---
title: "Big Spenders"
output:
  html_notebook: default
  html_document: default
---



##Introduction

In this short piece, I will investigate the most obvious question - Who spent the most?

-- Insert contents/explanation in Introduction
-- Insert background on what the data is, why am I doing this?
-- Make the piece self contained


##Libraries

TODO - Why these libraries?

_dplyr_

_readr_

_ggplot2_

```{r}
library(dplyr)
library(readr)
library(ggplot2)

```



##Data Load

I am using data from The Independent Parlimentary Standards Authority, who [provide data](http://www.parliamentary-standards.org.uk/DataDownloads.aspx) for all expense claims made by MP's, grouped by year.

This analysis is using the 2015 expense claims csv.

I'm able to set the type of some columns whilst reading in the data. Readr cannot individually set columns as factors though, so these are manually set following the load.


```{r}

#Read data to expenses 2015
expenses2015 <- read_csv("Data/DataDownload_2015.csv", 
                         col_types = cols(
                                        Date = col_date(format = "%d/%m/%Y"),
                                        `Claim No.` = col_character(),
                                        `Reason If Not Paid` = col_character(),
                                        `Amount Not Paid` = col_double()
                                         )
)


# Convert some columns to factor

#To do - Convert this to an elegant apply function...

# List of columns to convert to factors - `MP's Name`, `MP's Constituency`, Category,`Expense Type`, From, To, Travel

expenses2015$`MP's Name` <- as.factor(expenses2015$`MP's Name`)
expenses2015$`MP's Constituency` <- as.factor(expenses2015$`MP's Constituency`)
expenses2015$Category <- as.factor(expenses2015$Category)
expenses2015$`Expense Type` <- as.factor(expenses2015$`Expense Type`)
expenses2015$`MP's Name` <- as.factor(expenses2015$`MP's Name`)
expenses2015$From <- as.factor(expenses2015$From)
expenses2015$To <- as.factor(expenses2015$To)
expenses2015$Travel <- as.factor(expenses2015$Travel)
expenses2015$Status <- as.factor(expenses2015$Status)
expenses2015$`Reason If Not Paid` <- as.factor(expenses2015$`Reason If Not Paid`)



```



##Analysis
### 1 - Largest Expense Claims by Category

TODO - Add £ formatting to table and graph.
TODO - Neaten x-axis labels (remove and use legend?)

Using the `group_by` function from Hadley's `dplyr` package, we can group all our expense claims by category. Analysis of these groupings, such as mean, min, max, count and more advanced ideas can be explored.

To begin with, we will look at the maximum value claimed by an MP in each category. We may also wish to explore the mean, or median expense claim. A jitter plot could also be a useful way to explore the patterns in the data.

```{r}

expense_by_category <- expenses2015 %>% group_by(Category)

summarise(expense_by_category,  `Largest Claim (£)` = max(`Amount Claimed`), `Mean Claim (£)` = mean(`Amount Claimed`), `Median Claim (£)` = median(`Amount Claimed`), `Count of Claims` = n())

```

Whilst interesting, this isn't particularly insightful. Let's try visualising the data.

Things to note

* `geom_jitter` is the same as `geom_point` but with an inbuilt jitter, adding a slight element of random noise. This helps to visualise the data, spreading it out a little whilst maintaining the overall features.
* You can see that we actually have some negative claims plotted. These represent "refunded expenses" by the MP.

```{r}
ggplot(data = expenses2015, aes(y = `Amount Claimed`, x = Category, col = Category)) + geom_jitter(alpha = 0.3) + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none")
```
This is a much clearer way to display the data. We can immediately see  the rough range, maximum claim and general trend of the data.


### 2 - Largest Claims by Constituency

```{r}
expense_by_constituency <- expenses2015 %>% group_by(`MP's Constituency`)

summarise(expense_by_constituency,  `Largest Claim (£)` = max(`Amount Claimed`), `Mean Claim (£)` = mean(`Amount Claimed`), `Median Claim (£)` = median(`Amount Claimed`), `Count of Claims` = n())

```

#### 2.1 Going beyond constituency

We can gather additional information on MP Constituencies from [this wikipedia page using](https://en.wikipedia.org/wiki/List_of_United_Kingdom_Parliament_constituencies) `rvest`

I've saved the data to a local file `constituencies.txt` which we can work with.

We'll need to massage the data so that it joins correctly on `Constituency`

TO DO - Move data massaging to the appendix? Start with the prepped set?


```{r}

constits <- read_csv("Data/constituencies.txt")

constits$Constituency <- as.factor(constits$Constituency)
constits$`Largest ceremonial county or council area (Scotland)` <- as.factor(constits$`Largest ceremonial county or council area (Scotland)`)
constits$`Country of the UK` <- as.factor(constits$`Country of the UK`)
constits$Region <- as.factor(constits$Region)

levels(constits$Constituency)[648] <- "Ynys Mon"


# Use a fresh set of the data, avoiding complexities later on

expenses2015v2 <- expenses2015

names(expenses2015v2)[5] <- "Constituency"
levels(expenses2015v2$Constituency) <- sub(" [BC]C", "", levels(expenses2015v2$Constituency))
levels(expenses2015v2$Constituency)[645] <- "Ynys Mon"
View(expenses2015v2)

expenses2015v2 <- left_join(expenses2015v2, constits, by = "Constituency")


ggplot(data = expenses2015v2, aes(x = Region, y = `Amount Claimed`, col = Category), alpha = 0.2) + geom_jitter() + theme()


```

### 3 - Largest Overall Claims (by Value and by Total)


