---
title: "Exploratory Analysis"
author: "Connor Kirkpatrick"
date: "26 September 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(ggplot2)
library(knitr)
library(dplyr)

#Read in the 2015 Data

expenses2015 <- read_csv("Data/DataDownload_2015.csv", col_types = cols(
            Date = col_date(format = "%d/%m/%Y"),
            `Claim No.` = col_character(),
            `Reason If Not Paid` = col_character(),
            `Amount Not Paid` = col_double()
))


# Convert some columns to factor

#To do - Convert this to an elegant apply function...
# `MP's Name`, `MP's Constituency`, Category,`Expense Type`, From, To, Travel
expenses2015$`MP's Name` <- as.factor(expenses2015$`MP's Name`)
expenses2015$`MP's Constituency` <- as.factor(expenses2015$`MP's Constituency`)
expenses2015$Category <- as.factor(expenses2015$Category)
expenses2015$`Expense Type` <- as.factor(expenses2015$`Expense Type`)
expenses2015$`MP's Name` <- as.factor(expenses2015$`MP's Name`)
expenses2015$From <- as.factor(expenses2015$From)
expenses2015$To <- as.factor(expenses2015$To)
expenses2015$Travel <- as.factor(expenses2015$Travel)
expenses2015$Status <- as.factor(expenses2015$Status)
expenses2015$`Reason If Not Paid` <- as.factor(expenses2015$`Reason If Not Paid`)



```



```{r}


#What was the largest claim?
max(expenses2015$`Amount Claimed`)

# Did it match the largest amount paid?
max(expenses2015$`Amount Paid`)

#Vector of whether the amount claimed equals the amount paid
ac_to_ap <- expenses2015$`Amount Claimed` == expenses2015$`Amount Paid`

#ratio of rejected expenses
mean(ac_to_ap)
table(ac_to_ap)
rejected <- subset(expenses2015, !ac_to_ap)
mean(rejected$`Amount Claimed`)
str(rejected$`Reason If Not Paid`)

#reasons why they weren't paid
levels(rejected$`Reason If Not Paid`)

#Stats
table((rejected$`Reason If Not Paid`))

table(expenses2015$`Reason If Not Paid`)


ggplot(rejected) + geom_bar(aes(x = rejected$`Reason If Not Paid`, fill = rejected$`Reason If Not Paid` )) + labs(x = "Reason not paid", y = "Frequency", fill = NULL) + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none")


by_mp <- group_by(expenses2015, `MP's Name`)
total_by_mp <- summarise(by_mp, count = n(),
                total_claimed = sum(`Amount Claimed`))

ototal_by_mp <- arrange(total_by_mp, desc(total_claimed))

top10 <- slice(ototal_by_mp, 1:10)
top10

```



# Further Developments

* What do these MP's have in common?
* Who claimed the most dodgy expenses?
* Use genderizeR to predict Gender on the names?




## The basics


In this section, we will find the following statistics about our data:

* Max Expense Claimed
* Minimum Expense Claimed
* Mean Expense Claimed


To find the Max expense claimed, we can use the `max` function.

```{r}

maxclaimed <- max(expenses2015$`Amount Claimed`)
formatted_max_claimed <- paste("£", format(maxclaimed, digits=10, nsmall=2, decimal.mark=".", big.mark=","), sep = "")

whoclaimedmax <- which.max(expenses2015$`Amount Claimed`)

most_expensive_claim <- expenses2015[whoclaimedmax,]

kable(most_expensive_claim, format = "markdown")


```

So we can see that `r most_expensive_claim[4]` had the largest claim (of `r formatted_max_claimed`!). We can see this was for "`r most_expensive_claim[7]`"


Let's now do the same for the smallest expense claimed.

```{r}

minclaimed <- min(expenses2015$`Amount Claimed`)
who_claimed_min <- which.min(expenses2015$`Amount Claimed`)
#least_expensive_claim <- expenses2015[who_claimed_min] commented out as it is broken
```


It seems we've hit an error. The smallest amount claimed is in fact a negative number! We should investigate to see what is happening here.

```{r}

hist(expenses2015$`Amount Claimed`)
table(expenses2015$`Amount Claimed` < 0)

```

So there are currently 1874 expenses which have an `Amount Claimed` value of less than 0. Let's filter for these and have a deeper look.


```{r}

negative_claims <- filter(expenses2015, `Amount Claimed` < 0)
kable(head(select(negative_claims, c(2:5, 16:17))))


```


Looking at the claims, we can see that they look like genuine expenses, but with a negative sign in front of them. So possible explanations could be:

* Error in the data. To fix, we could just remove the "-" sign.
* These may be correct expenses. We'd need to investigate the context for these further if that was the case.