---
title: "Exploratory Analysis"
author: "Connor Kirkpatrick"
date: "26 September 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(ggplot2)
library(knitr)
library(dplyr)
library(lubridate)

#Read in the 2015 Data

expenses2015 <- read_csv("Data/DataDownload_2015.csv", col_types = cols(
            Date = col_date(format = "%d/%m/%Y"),
            `Claim No.` = col_character(),
            `Reason If Not Paid` = col_character(),
            `Amount Not Paid` = col_double()
))


# Convert some columns to factor

#To do - Convert this to an elegant apply function...
# `MP's Name`, `MP's Constituency`, Category,`Expense Type`, From, To, Travel
expenses2015$`MP's Name` <- as.factor(expenses2015$`MP's Name`)
expenses2015$`MP's Constituency` <- as.factor(expenses2015$`MP's Constituency`)
expenses2015$Category <- as.factor(expenses2015$Category)
expenses2015$`Expense Type` <- as.factor(expenses2015$`Expense Type`)
expenses2015$`MP's Name` <- as.factor(expenses2015$`MP's Name`)
expenses2015$From <- as.factor(expenses2015$From)
expenses2015$To <- as.factor(expenses2015$To)
expenses2015$Travel <- as.factor(expenses2015$Travel)
expenses2015$Status <- as.factor(expenses2015$Status)
expenses2015$`Reason If Not Paid` <- as.factor(expenses2015$`Reason If Not Paid`)



```



```{r}


#What was the largest claim?
max(expenses2015$`Amount Claimed`)

# Did it match the largest amount paid?
max(expenses2015$`Amount Paid`)

#Vector of whether the amount claimed equals the amount paid
ac_to_ap <- expenses2015$`Amount Claimed` == expenses2015$`Amount Paid`

#ratio of rejected expenses
mean(ac_to_ap)
table(ac_to_ap)
rejected <- subset(expenses2015, !ac_to_ap)
mean(rejected$`Amount Claimed`)


#reasons why they weren't paid
levels(rejected$`Reason If Not Paid`)

#Stats
table((rejected$`Reason If Not Paid`))

table(expenses2015$`Reason If Not Paid`)


ggplot(rejected) + geom_bar(aes(x = rejected$`Reason If Not Paid`, fill = rejected$`Reason If Not Paid` )) + labs(x = "Reason not paid", y = "Frequency", fill = NULL) + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none")


by_mp <- group_by(expenses2015, `MP's Name`)
total_by_mp <- summarise(by_mp, count = n(),
                total_claimed = sum(`Amount Claimed`))

ototal_by_mp <- arrange(total_by_mp, desc(total_claimed))

top10 <- slice(ototal_by_mp, 1:10)
kable(top10, format = "markdown")

```



# Further Developments

* What do these MP's have in common?
* Who claimed the most dodgy expenses?
* Use genderizeR to predict Gender on the names?




## The basics


In this section, we will find the following statistics about our data:

* Max Expense Claimed
* Minimum Expense Claimed
* Mean Expense Claimed


To find the Max expense claimed, we can use the `max` function.


Before we do that however, let's make a quick helper function to print the numbers nicely as a currency

Note that an alternative way (probably better) would be to create a "currency" class, and then creat a print.currency method that would automatically be called when our currency class was printed.

```{r}



#Convert to Pounds

# Takes a numeric and returns a string of the value in the format £X,XXX.XX

con_to_pounds <- function(value, pence = T){
    if(pence == T){
        n_small <-  2
    }else{
        n_small <-  0
    }
    paste("£", format(value, digits=10, nsmall=n_small, decimal.mark=".", big.mark=","), sep = "")
}
```

Now that we have our helper function, let's try it out.

```{r max_claim}

max_claimed <- max(expenses2015$`Amount Claimed`)

formatted_max_claimed <- con_to_pounds(max_claimed)

whoclaimedmax <- which.max(expenses2015$`Amount Claimed`)

most_expensive_claim <- expenses2015[whoclaimedmax,]

kable(most_expensive_claim, format = "markdown")


```

So we can see that `r most_expensive_claim[4]` had the largest claim (of `r formatted_max_claimed`!). We can see this was for "`r most_expensive_claim[7]`"


Let's now do the same for the smallest expense claimed.

```{r min_claim}

minclaimed <- min(expenses2015$`Amount Claimed`)
who_claimed_min <- which.min(expenses2015$`Amount Claimed`)
#least_expensive_claim <- expenses2015[who_claimed_min] commented out as it is broken
```


It seems we've hit an error. The smallest amount claimed is in fact a negative number! We should investigate to see what is happening here.

```{r}

hist(expenses2015$`Amount Claimed`)
table(expenses2015$`Amount Claimed` < 0)

```

So there are currently 1874 expenses which have an `Amount Claimed` value of less than 0. Let's filter for these and have a deeper look.


```{r}

negative_claims <- filter(expenses2015, `Amount Claimed` < 0)
kable(head(select(negative_claims, c(2:5, 16:17))))


```


Looking at the claims, we can see that they look like genuine expenses, but with a negative sign in front of them. So possible explanations could be:

* Error in the data. To fix, we could just remove the "-" sign.
* These may be correct expenses. We'd need to investigate the context for these further if that was the case.


Having emailed IPSA (who provded the data), it appears that these negative expenses arise when an MP has been refunded for an expense claimed, and so in turn has the expense refunded.

An easy solution would be to filter out these negative values, as they're not relevant to what we're looking at currently.

* We could investigate the refunded expenses to see if there is a pattern to them that differs from normal expenses.
* We could in fact use these expenses to find their original claim, and then remove both.


For now, let's just keep it simple and filter out the negative values

```{r}

expenses2015postive <- filter(expenses2015, `Amount Claimed` > 0)

#Alternatively for our purposes, we could have just used expenses2015[expenses2015$`Amount Claimed` >= 0]
```

Now let's try again to find the smallest claim made

```{r}

minclaimed <- min(expenses2015postive$`Amount Claimed`)
who_claimed_min <- which.min(expenses2015postive$`Amount Claimed`)

least_expensive_claim <- expenses2015postive[who_claimed_min,]


kable(least_expensive_claim)


```

## The average claim?

In this section, we will find the value of the average claim. We'll also find the top 5 Expense types for approved and non approved claims.

### Mean Claim and Mean Paid
```{r}
mean_claimed <- mean(expenses2015postive$`Amount Claimed`)
mean_paid <- mean(expenses2015postive$`Amount Paid`)

con_to_pounds(mean_claimed) # con_to_pounds clearly isn't perfect, as it's printing too many sig figs
con_to_pounds(mean_paid) # same here


```

### Most common expense categories


```{r}

common_expense_types <- expenses2015postive %>%
                        group_by(Category) %>% 
                        dplyr::summarise(count = n(),
                                         sum = con_to_pounds(sum(`Amount Claimed`)),
                                        mean_claim = con_to_pounds(mean(`Amount Claimed`)))


kable(arrange(common_expense_types, desc(count))[1:5,])

ggplot(common_expense_types) + geom_bar(aes(x = Category, y = sum, fill = Category), stat = "identity") +labs(x = "Category of Expense", y = "Sum Claimed", fill = NULL) + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none")

```


Whilst `r arrange(common_expense_types, desc(count))[1,1]` is the most common claim category, you can see that by total sum and mean claim, `r arrange(common_expense_types, desc(count))[1,1]`



```{r}

# Convert this to filter for rejected claims. Search for categories/types of reject reasons

common_rejected_expense_types <- expenses2015 %>%
                                 filter(Status == "Not Paid") %>%
                                  group_by(Category) %>% 
                                  dplyr::summarise(count = n(),
                                                 sum = con_to_pounds(sum(`Amount Claimed`)),
                                                mean_claim = con_to_pounds(mean(`Amount Claimed`)))


kable(arrange(common_rejected_expense_types, desc(count))[1:5,])

ggplot(common_rejected_expense_types) + geom_bar(aes(x = Category, y = sum, fill = Category), stat = "identity") +labs(x = "Category of Expense", y = "Sum Claimed", fill = NULL) + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none")

```


# Time Analysis 

##Does the average expense claim vary across the months of the Year?


To do this, we will use the `lubridate` library (good thing we loaded it earlier!).


```{r expenses_by_month}

expenses_by_month <- expenses2015 %>%
                        filter(`Amount Claimed` > 0) %>%
                        group_by(month(Date))
  

summary_expenses_by_month <- summarise(expenses_by_month , `Number of Claims` = n(), `Sum of Claims` = con_to_pounds(sum(`Amount Claimed`)))

kable(summary_expenses_by_month)


ggplot(expenses_by_month) + geom_bar(aes(x = month(Date), fill = Category)) + labs(x = "Month", y = "Number of Claims", title = "Bar graph to show the number of claims\n submitted by month")

```

Things to notice

* The number of accomodation claims stays roughly constant across the month, with a slight peak in June, and a dip in August.
* MP's Travel claims have a huge decease in April (and a slighlty less dramatic dip in August)
* Claims for winding up are submitted between May and August.


Let's make the same graph, but displaying the sum of the claims instead.

```{r}

## This graph is not correct!

ggplot(expenses_by_month) + geom_bar(aes(x = `month(Date)`, weight = `Amount Claimed`, fill = Category)) + labs(x = "Month", y = "Sum of Claims", title = "Bar graph to show the sum of claims\n submitted by month")

                                     